---
title: "Homework 10"
output: pdf_document
date: "2024-12-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

a.

```{r}
require(faraway)
require(ggplot2)
require(dplyr)

data(rats)

rats %>% 
  ggplot(aes(x = poison, y = time)) +
  geom_point() +
  stat_summary(fun.y = "mean", geom = "line", aes(group = treat, color = treat)) +
  theme(legend.position = "top", legend.direction = "horizontal")

rats %>% 
  ggplot(aes(x = treat, y = time, shape = poison)) +
  geom_point() +
  stat_summary(fun.y = "mean", geom = "line",
               aes(group = poison, color = poison)) +
  theme(legend.position = "top", legend.direction = "horizontal")


```

b. The interaction effect is not significant. 

```{r}

lmod <- lm(time ~ poison*treat, rats)
anova(lmod)

```

c. There does appear to be statistical significance between poisons and between treatments. 

```{r}

anova(lm(time ~ treat + poison, rats))

```

d. Diagnostics have not been met and give troublesome results. The Q-Q plot has long-tailed errors, and the plot of fitted values against residuals shows some concerning symmetry. 

```{r}

qqnorm(residuals(lmod), main="")
qqline(residuals(lmod)) ## long-tailed errors

plot(fitted(lmod), residuals(lmod), xlab = "Fitted",
     ylab = "Residuals") ## symmetry
abline(h=0)

plot(residuals(lmod) ~ poison, rats, ylab = "Residuals")


## Test for difference in variance
ratse <- rats[(1:48),]
ratse$res <- sqrt(abs(residuals(lmod))[(1:48)])
vmod <- lm(res ~ poison + treat, ratse)
anova(vmod)

```

e. Looking at the plot from question 1 and the model without interaction, it is hard to say that one treatment is *clearly* superior to the others. Treatment B appears to be the most effective, but treatment D is close. And for poison III, they are about the same. I would say that treatment B appears to be the best treatment but that it is not clearly superior to D. 

```{r}

rats %>% 
  ggplot(aes(x = poison, y = time)) +
  geom_point() +
  stat_summary(fun.y = "mean", geom = "line", aes(group = treat, color = treat)) +
  theme(legend.position = "top", legend.direction = "horizontal")

summary(lm(time ~ treat + poison, rats))

```

f. We find lambda to be -0.82828 and after transforming the data see better results. We see higher significance in poison II and treatments C and D. We also notice that treatment B and D had large changes to their coefficients, as did poison III. 

```{r}

require(MASS)
bc <- boxcox(lmod)
lambda <- bc$x[which.max(bc$y)]

time_new <- (rats$time^lambda - 1) / lambda

lmod_main <- lm(time ~ poison + treat, rats)
lmod_transformed <- lm(time_new ~ poison + treat, rats)

summary(lmod_main)
summary(lmod_transformed)

qqnorm(residuals(lmod_transformed))
qqline(residuals(lmod_transformed))

```

## 2. 

We find the width of the band to be 31.10935 using the qtukey() function. Looking at a table of differences, it would appear that the differences between B and E, B and F, B and G, and B and H are significantly different at the 5% level. We also see that C and E, C and F, C and G, C and H, D and F, D and G, and D and H are significantly different. 

```{r}

data(OrchardSprays)

xtabs(decrease ~ rowpos + colpos, OrchardSprays)
matrix(OrchardSprays$treatment,8,8,byrow = TRUE)

lmodLatin <- lm(decrease ~ treatment + as.factor(rowpos) + as.factor(colpos), OrchardSprays)
drop1(lmodLatin, test="F")
summary(lmodLatin)
qtukey(0.95, 8, 42)*9.757/sqrt(2) ## 8 is no. of treatments, 42 deg of freedom

scoefs <- c(0, coef(lmodLatin)[2:8])
outer(scoefs, scoefs, "-")

```