---
title: "Homework 2"
author: "Joe Dickerson"
date: "2025-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Packages
library(faraway)
library(tidyverse)

data(seeds)

```

3.

(a) It appears that when the container is not covered, we see germination increase as moisture increases. With a cover, we see higher germination at the lower moisture levels, and lower germination at the higher moisture levels. It appears that the cover helped to create an ideal environment for the seeds at lower moisture levels than when the seeds are uncovered. 

```{r}

seeds %>% 
  ggplot(aes(y = germ, x = moisture)) +
  geom_point() +
  facet_grid( ~ covered) +
  labs(x = "Moisture level", y = "Germination Percentage") 

```

(b) There does not appear to be a box effect based on the plot.

```{r}

df <- seeds %>% 
  mutate(box = rep(1:8, each = 6))

df %>% ggplot(aes(y = germ, x = moisture, color = factor(box))) +
  geom_point() +
  geom_line() +
  facet_grid( ~ covered) +
  labs(x = "Moisture level", y = "Germination Percentage")

```

(c)

```{r}

glmod <- glm(cbind(germ, 100 - germ) ~ moisture + covered + factor(box), family = binomial, df)

summary(glmod)
glmod_sum = summary(glmod)


```
(d) We have very low p-values, which would indicate problems with the model.

```{r}

pchisq(deviance(glmod), df.residual(glmod), lower = F)

pearsonstat = sum(residuals(glmod, "pearson")^2)

pchisq(pearsonstat, glmod_sum$df.residual, lower.tail = F)

```

(e)

```{r}

predict(glmod, moisture = 13)

```

(f) Our fitted values against residuals plot shows a nonlinear pattern. 

```{r}

glmfitted <- fitted(glmod)
glmresiduals <- residuals(glmod)

plot(residuals(glmod) ~ fitted(glmod))
abline(h=0)

```

(g)

```{r}

## removed the 47th row to fit with the residuals as the original dataset is missing a value here. The model omits this row

df_new <- df[-47,]

df_new %>% ggplot(aes(x = moisture, y = glmresiduals, color = covered)) +
  geom_point() +
  geom_smooth(fill=NA)

```

(h) 

```{r}

hatv <- hatvalues(glmod)
halfnorm(hatv, ylab = "Leverages")

qqnorm(residuals(glmod))
qqline(residuals(glmod))

```