---
title: "Homework 3"
author: "Joe Dickerson"
date: "2025-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

## packages

library(faraway)
library(tidyverse)
library(MASS)
library(survival)

```

3. 

(a)

It does appear that there is a higher likelihood of spontaneous abortions among the case group when keeping in mind there there are two observations of control for every one case observation. For example, 22 instances of a single spontaneous abortion with 0 induced in the case group, and 25 in the control. With half the observations in the case group, this is a higher rate. Looking at the same column of 0 induced abortions, we have similar findings with the count of two spontaneous abortions, 18 in the case group and 11 in the control. This is a much higher rate in the case group than the control. 

```{r}

data(infert)

case <- infert[infert$case == 1,]
control <- infert[infert$case == 0,]



case_table <- table(case$spontaneous, case$induced)
case_table <- addmargins(case_table)
control_table <- table(control$spontaneous, control$induced)
control_table <- addmargins(control_table)

rownames(case_table) <- paste("Spontaneous", rownames(case_table))
colnames(case_table) <- paste("Induced", colnames(case_table))
rownames(control_table) <- paste("Spontaneous", rownames(control_table))
colnames(control_table) <- paste("Induced", colnames(control_table))

cat("Cases")
case_table

cat("\nControls")
control_table

```

(b) 

We have an odds ratio of 3.31 for spontaneous and 1.52 for induced, which suggests higher odds for a case (infertile) when a spontaneous abortion has occurred than an induced one. Both values are greater than one, so both increase the odds of a case. 


```{r}

glmod <- glm(case ~ spontaneous + induced, family = binomial, infert)

summary(glmod)

exp(coef(glmod))


```

(c) 

Without spontaneous and induced, we are only measuring the effect of education and age on the likelihood of a person being infertile. It would be reasonable to suggest that these factors alone would not have a great impact on the response. However, the contribute to the matching of case vs control.

```{r}

glmod2 <- glm(case ~ education + age + parity, family = binomial, infert)

summary(glmod2)

```

(d)  

With all five predictors in the model, there seems to be a higher odds ratio for spontaneous and induced than without the additional predictors. We also see values less than 1 for education and parity. Age is 1.04, which is very close to 1.

```{r}

glmod3 <- glm(case ~ spontaneous + induced + education + age + parity, family = binomial, infert)

summary(glmod3)

exp(coef(glmod3))

```

(e) 

We see similar results in the summary for spontaneous and induced, as well as our odds ratio. It seems that if a person has had a spontaneous abortion, the odds are higher than they will be infertile. The same is true for an induced abortion. 

```{r}

cmod = clogit(case ~ spontaneous + induced + education + age + parity + strata(stratum), infert)

summary(cmod)

exp(coef(cmod))

```

(f) 

The significance of spontaneousOrd.L and inducedOrd.L seem to indicate that there is in fact a linear relationship in the predictors, but not a quadratic one due to the insignificance of spontaneousOrd.Q and inducedOrd.Q. It appears there is could be an advantage to making the distinction.

```{r}

infert$spontaneousOrd <-ordered(infert$spontaneous, levels = 0:2, labels = c("0", "1","2+"))

infert$inducedOrd <-ordered(infert$induced, levels = 0:2, labels = c("0", "1","2+"))

cmodr = clogit(case ~ spontaneousOrd + inducedOrd + education + age + parity + strata(stratum), infert)

summary(cmodr)

exp(coef(cmodr))

```