---
title: "R Exercises (05)"
author: "Joe Dickerson"
date: "2025-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.

```{r}

library(tidyverse)
library(NHANES)

data(NHANES)

```

2. 

```{r}

NHANES %>% 
  ggplot(aes(x = BMI, y = TotChol)) +
  geom_point()

```

3. Setting the color in the aesthetic here does not change the color of the geom_point layer, but it is trying to stratify the data points in the plot by a string, "blue" in this case, which is why we see the "blue" in the legend. It would have to go outside of the aes() somewhere.
```{r}

ggplot(data = NHANES) + 
  geom_point(mapping = aes(x = Height, y = Weight, color = "blue"))

```

4. I am unable to map a continuous variable to shape because shape can only take categorical variables. 

```{r}

NHANES %>% 
  ggplot(aes(x = Height, y = Weight, color = Poverty, size = Poverty, shape = Education)) +
  geom_point()

NHANES %>% 
  ggplot(aes(x = Height, y = Weight, shape = Education)) +
  geom_point()

```

5. The stroke aesthetic appears to modify the width of the point layer for a few specific point shapes (per the geom_point documentation).

```{r}

NHANES %>% 
  ggplot(aes(x = Height, y = Weight)) +
  geom_point(shape = 23, size = 3, color = "blue", fill = "green", stroke = 3) 

```

6.

```{r}

NHANES %>% 
  ggplot(aes(x = Height, y = Weight, color = interaction(Gender,Race1))) +
  geom_line()

```

7.

```{r}

NHANES %>% 
  ggplot(aes(x = Height, y = Weight)) +
  geom_line() + 
  facet_grid(~ Gender)

NHANES %>% 
  ggplot(aes(x = Height, y = Weight)) +
  geom_line() + 
  facet_grid(~ Race1)

```

8.

```{r}

NHANES %>% 
  ggplot(aes(x = BPSysAve, y = TotChol, color = Gender)) +
  geom_point() +
  geom_smooth(color = "black", size = .5) +
  facet_grid(SmokeNow ~ Gender) +
  labs(y = "Total cholesterol (mg/dl)", x = "Blood pressure (mm Hg)", title = "Blood Pressure vs. Gender vs. Smoking Status", caption = "Source: NHANES survey 2009-2012")

```