---
title: 
author: 
subtitle: 
date: 
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

## you may put all R packages here.
suppressWarnings(suppressMessages(library(tidyverse)))
require(ggplot2)

```


#### Dataset

The prepared data file is available at
https://Caleb-Huo.github.io/teaching/data/pressData/benchPress.csv.
**Your analysis should base on this data**.

#### Dataset characteristics

- ID: participant id
- time: in terms of days (since entry into the study) for bench press record
- press: the amount of bench press (in lbs)
- weight: the body weight of the participant at the beginning of study
- age: the age of the participant at the beginning of the study
- prev: indicator for whether the participant has the previous weight lifting experience (0: no; 1: yes)
- diet: indicator for whether the participant is not in the diet program (0) or in the diet program (1)



## Analysis

```{r}

data <- read_csv("https://Caleb-Huo.github.io/teaching/data/pressData/benchPress.csv")

```


### Question 1. (10 pts)
How many participants are with age <= 20 and in the diet program (i.e., diet == 1)?
Among them, what is the percentage of prior exercise experience (i.e., pre == 1)?

```{r}

x <- data %>% 
  filter(diet == 1 & age <= 20) %>% 
  group_by(ID) %>% 
  count(ID) %>% 
  pull(ID) %>% 
  length()

paste(x,"participants")

y <- data %>% 
  filter(diet == 1 & age <= 20 & pre ==1) %>% 
  group_by(ID) %>% 
  count(ID) %>% 
  pull(ID) %>% 
  length()

paste0(round((y/x) * 100,2),"% with prior experience")

```


### Question 2. (10 pts)
Prepare a demographic tables by dietary program. 
The result should look like the table below, but please use R and kable function of the knitr packages to prepare this table.
For age (continuous variable), calculate its mean and standard deviation for each group;
for previous experience (prev==1; binary variable), calculate its counts and percentage for each group.


| Type            |   diet==0  | diet==1 | 
|-----------------|-----------|---------------|---------------------------|
| `N`             |    count       | count                     |
| `Age (years)`   |   mean (SD)   | mean (SD)                 |
| `Previous experience == 1, n (%)`   | count (%) |   count (%)   |



```{r}

require(knitr)
require(kableExtra)

dataDemographics <- data %>% 
  group_by(ID) %>% 
  summarise(ID = ID[1],
         age = age[1],
         diet = diet[1],
         pre = pre[1]) %>% 
  group_by(diet) %>%
  summarise(count = n(),
            mean_age = mean(age),
            sd_age = sd(age),
            pre_count = sum(pre),
            pre_percent = mean(pre) * 100
            )

## Used to acquire data for the table:

# round(dataDemographics,2)


newtable <- data.frame(
  "Type" = c("`N`", "`Age (years)`", "`Previous experience == 1, n(%)`"),
  "diet==0" = c("50", "22.08 (2.51)", "13 (26%)"),
  "diet==1" = c("50", "21.88 (2.93)", "14 (28%)")
)

kable(newtable,
      col.names = c("Type", "diet==0", "diet==1"), 
      ) %>% 
  kable_styling() %>%
  column_spec(1) %>%
  row_spec(0, align = "c")
  

```





### Question 3. (10 pts)
Write a function, the input is a vector. 
The output is a list of max, min,interquartile range (IQR) of the vector.
If the input vector contains any negative values, provide a warning message.
Then apply this function to a vector which represents the baseline weight of all participants.

```{r}

fncVEC <- function(x){
  if(any(x < 0)) {
    warning("The input contains a negative value.")
  }
  
  max1 <- max(x)
  min1 <- min(x)
  iqr1 <- IQR(x)
  
  output <- list(
    Maximum = max1,
    Minimum = min1,
    IQR = iqr1
  )
  
  return(output)
}

weightvec <- as.vector(data$weight)

fncVEC(weightvec)

```

### Question 4. (10 pts)
Sort the data by age (ascending), ID (descending) and time (descending). 
Then print out the first five instances of the amount of bench press (i.e., press)

```{r}

sortedData <- data %>% 
  arrange(age, desc(ID), desc(time)) %>% 
  slice(1:5)

sortedData$press

```




### Question 5. (10 pts)

Draw histograms for weight at baseline of all participants, 
stratified by each combination of pre and diet (there should be 4 panels in the figure).
The height of the histogram should represent the density instead of the frequency.
Also draw the density function (in red color) on top of the histogram.

```{r}

diet.labs <- c("Not in Diet Program", "In Diet Program")
names(diet.labs) <- c("0", "1")

pre.labs <- c("No Prior Lifting Experience", "Prior Lifting Experience")
names(pre.labs) <- c("0", "1")

data %>% 
  ggplot(aes(x=weight, y=..density..)) +
  geom_histogram(bins = 20) +
  geom_density(color = "red") +
  facet_grid(pre ~ diet, labeller = labeller(diet = diet.labs, pre = pre.labs))
  
  #facet_grid(cols = vars(diet), rows = vars(pre))

```


### Question 6. (10 pts)
We want to do some string manipulations for the attribute **time**.
For each of the 839 instances, 
Change 1,3,5,7,9 to O (Odd); change 2,4,6,8,0 to E (Even).
For example, 154 should become "OOE"; 166 should become "OEE"; 24 should become "EE"; 0 should become "E";
How many instances have more letter E than letter O? (i.e., #E > #O)

```{r}

dataSTR <- data 

## Function for replacing the digits with the letters
fncSTR <- function(num){
  digits <- strsplit(as.character(num), "")[[1]]
  replaced <- ifelse(as.integer(digits) %% 2==0, "E", "O")
  paste(replaced, collapse = "")
}

## Applying the function to the "time" column
dataSTR0 <- dataSTR %>% 
  mutate(time = sapply(time, fncSTR))


## Function to count the instances 
fncLOG <- dataSTR0 %>% 
  filter(sapply(time, function(x){
    sum(strsplit(x,"")[[1]] == "E") > sum(strsplit(x,"")[[1]] == "O")
  })) %>% 
  nrow()

fncLOG

```



### Question 7. (10 pts)
Plot the individual trajectory of the bench pressed weights (lbs) over time for (a) men in the no diet group and (b) men in the diet group. 
So there will be two sub-figure, each contains 50 trajectories from each group. 
Do not include a figure legend, which may make the figure messy.

```{r}

data %>% 
  ggplot(
  aes(x = time, y = press, group = ID, color = factor(diet))
) + 
  geom_line(alpha = .5) +
  facet_wrap(. ~ diet) +
  theme(legend.position="none") +
  labs(title = "Trajectory of Weight vs Time")

```




### Question 8. (10 pts)

For each subject, define the improvement of bench press as the difference of the bench press weight at the final visit and the initial bench press weight (e.g., $press_{final} - press_{initial}$).
We want to visualize the the improvement of bench press using box plot and jitter plot, stratified by different dietary programs (diet).

```{r}

dataImprove <- data %>% 
  group_by(ID) %>% 
  summarise(Improvement = last(press)-first(press),
            Diet_Program = max(diet)
            ) %>% 
  mutate(
    "Dietary Program" = ifelse(Diet_Program == 1, "In the Diet Program", "Not in the Diet Program" )
  )

ggplot(dataImprove) +
  aes(x=`Dietary Program`, y=Improvement, color = `Dietary Program`) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  labs(title = "Bench Press Improvement", color = "Dietary Programs", x="") +
  theme(legend.position="none")

```






### Question 9. (10 pts)
We want to visualize the mean trajectories over time for both diet group and no diet group.
Since the times at which bench press was recorded for each man varied across men,
we will only plot three time points: day 0, day 150, and day 300.
If there are no days that match exactly day 150,
**we can use the time closest to day 150 to replace**.
We will use the bench press weight of the last observation as the bench press weight for day 300.
Use different colors to indicate diet/no diet groups.
Mark standard error bars at the mean estimates of all these three time points.
Everything should be within one figure and one panel.

```{r, message=FALSE}

dataMean <- data %>% 
  mutate(time150 = abs(time - 150)) %>% 
  group_by(ID, diet) %>% 
  summarise(
    Press0 = press[1],
    Press150 = press[which.min(time150)],
    Press300 = press[which.max(time)]
  ) %>% 
  group_by(diet) %>% 
  summarise(
    Mean0 = mean(Press0),
    Mean150 = mean(Press150),
    Mean300 = mean(Press300)
  )

dataPlot <- dataMean %>% 
  pivot_longer(cols = -diet) %>% 
  mutate(time = ifelse(name=="Mean0", 0,
                       ifelse(name=="Mean150", 150, 300
                              )))

dataPlot %>% 
  ggplot(aes(x=time, y=value, group=diet, color=factor(diet))) +
  geom_line()
#geom_errorbar()

```


